Transmembrane helices are also an overlooked source of major histocompatibility complex class II epitopes and evolutionary more conserved than expected by change

### 2019-03-08 article.tex
Article is getting there

% Affiliations
\usepackage{authblk}
\title{pirouette: Determining the error BEAST2 makes in inferring a phylogeny}
\author[1]{Rich\`el J.C. Bilderbeek}
\author[1]{?Rampal S. Etienne}
\author[2]{Frans Bianchi}
\affil[1]{Groningen Institute for Evolutionary Life Sciences, University of Groningen, Groningen, The Netherlands}
\affil[2]{Frans' Institute, University of Groningen, Groningen, The Netherlands}

@@

About a fourth of the human proteome is anchored by transmembrane helices (TMHs) to
lipid membranes. It, therefore, comes as no surprise that
bioinformatics analysis predicts an over-representation of TMHs among strong MHC
class I (MHC-I) binders. Published peptide elution studies confirm that TMHs are indeed
presented by MHC-I. This raises the question how membrane proteins are processed
for MHC-I (cross-)presentation, with current research focusing on soluble antigens. The
presentation of membrane-buried peptides is likely important in health and disease, as
TMHs are considerably conserved and their presentation might prevent escape muta-
tions by pathogens. Therefore, it could contribute to the disease correlations described
for many human leukocyte antigen haplotypes.

\end{abstract}

{\bf Keywords:} antigen presentation, antigen cross-presentation, membrane proteins, bioinformatics, adaptive
immunity, transmembrane domain, epitopes, T lymphocyte, MHC-2

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\cite{bianchi2017transmembrane}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Methods}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

There are thousands of MHC2 alleles known.
For this research, we picked the three alleles that are most abundant,
accoring to \richel{reference here}, 
which are DRB4*0101, 
HLA-DPA10103-DPB10402 
\richel{
  we decided for DPB1*0402, but there are multiple alleles matching.
  I just picked the first one of the alleles matching, which are:
  HLA-DPA10103-DPB10402,
  HLA-DPA10104-DPB10402,
  HLA-DPA10105-DPB10402,
  HLA-DPA10106-DPB10402,
  HLA-DPA10107-DPB10402,
  HLA-DPA10108-DPB10402,
  HLA-DPA10109-DPB10402,
  HLA-DPA10110-DPB10402,
  HLA-DPA10201-DPB10402,
  HLA-DPA10202-DPB10402,
  HLA-DPA10203-DPB10402,
  HLA-DPA10204-DPB10402,
  HLA-DPA10301-DPB10402,
  HLA-DPA10302-DPB10402,
  HLA-DPA10303-DPB10402,
  HLA-DPA10401-DPB10402
}
and DQA1*0501/DQB1*0301,
that are present in respectively 31, 41 and 59 percent 
\richel{fix numbers}
of the current human population.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Results}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\input{table_imoimo.latex}
% has label tab:results

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusion}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Discussion}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Acknowledgements}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Data Accessibility}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Authors' contributions}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Bibliography
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% MEE style
\bibliographystyle{mee}
\bibliography{article}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Appendix section 1}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

>>>>

% Affiliations
\usepackage{authblk}
\title{
  Transmembrane helices are also 
  an overlooked source of major histocompatibility complex class II epitopes
  and evolutionary more conserved than expected by change
}
\author[1]{Rich\`el J.C. Bilderbeek}
\author[1]{?Rampal S. Etienne}
\author[2]{Frans Bianchi}
\affil[1]{Groningen Institute for Evolutionary Life Sciences, University of Groningen, Groningen, The Netherlands}
\affil[2]{Frans' Institute, University of Groningen, Groningen, The Netherlands}

@@

Transmembrane helices (TMHs) are an overlooked source of 
major histocompatibility complex (MHC) class I epitopes.
It is unknown why this is the case and 
whether MHC-II also is an overlooked source of epitopes.
This study investigates if there is an evolutionary conservation 
of a pathogen's TMHs and if this is matches 
with MHC-II's over/equally/under-presenting TMHs.
We find that ...
\end{abstract}

{\bf Keywords:} antigen presentation, membrane proteins, bioinformatics, 
adaptive immunity, transmembrane domain, epitopes, T lymphocyte, MHC-2

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

For MHC-1, we know that epitopes derived from transmembrane helices (TMHs) are 
over-presented by all human leukocyte antigen (HLA)-A and most HLA-B super 
types [\cite{bianchi2017transmembrane}].

A hypothesis offered by that same paper is that the presentation of 
especially TMHs may have an evolutionary advantage for the (human) host.
The argumentation given is that the presentation of parts of TMHs
is likelier to detect pathogenic proteins, as TMHs are already conserved
due to the functional requirement of being able to span a membrane.
Additionally, there are fewer escape mutations possible for the pathogen,
as relatively less of these will result in a functional TMH.

However, that argumentation can also go the other way: as there are less
polypeptide fragments from (the less amount of) possible TMHs, 
it is likelier that host and pathogen
share the same protein sequence, 
resulting the pathogen avoiding detection.

Within the immune response, MHC-I is unique in [doing something].
The other most important participant in the immune response in MHC-II,
which is more involved in detecting bacterial pathogens, 
e.g. tuberculosis, [and others].

From the earlier research done on MHC-I, it naturally follows to check
if MHC-II are likelier to present TMHs than expected by chance.

Extending the previous study on MHC-I, we also measure the evolutionary
conservation of TMH epitopes in Mycoplasma pathogens.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Methods}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Transmembrane helices and strong MHC-II-binding peptides
were predicted for a tuberculosis reference proteome 
from \url{https://www.ebi.ac.uk/reference_proteomes}
using the \verb;epitopeome; 
\richel{
  I'd enjoy a better name
}
R package [\cite{epitopeome}].
We picked the three MHC-II alleles that are most abundant 
in the current human population, 
which are DRB4*0101, HLA-DPA10103-DPB10402 
\richel{
  We decided for DPB1*0402, but there are multiple alleles matching.
  I just picked the first one of the alleles matching, which are:
  HLA-DPA10x-DPB10402, where x equals 103-110, 201-204, 301-303 or 401
}
and DQA1*0501/DQB1*0301 \richel{reference here}.
The 5\% peptides with the lowest IC50 values were defined as binders.
We then simply counted the number of amino acids that were present inside the
cell, within the membrane or outside of the cell, as well as if it was part 
of a strong MHC-II binding site, as shown in figue \ref{tab:results}.

The \verb;epitopeome; R package \cite{epitopeome} binds together 
the \verb;tmhmm; [\cite{tmhmm}] and \verb;netmhc2pan; [\cite{netmhc2pan}] 
R packages. \verb;tmhmm; provides an R interface to
TMHMM [\cite{krogh2001predicting}, \cite{sonnhammer1998hidden}], a tool
to predict where membrane proteins' amino acids are located within the
membrance.   
\verb;netmhc2pan; provides an R interface to
NetMHC2pan [\cite{jensen2018improved}], a tool
to predict MHC-II binding to proteins.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Results}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\input{table_imoimo.latex}
% has label tab:results
\richel{
  I'd enjoy 
  (1) a row with 'expected by chance', 
  (2) using percentages instead,
  (3) merging inside and outside
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusion}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Discussion}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Acknowledgements}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Data Accessibility}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Authors' contributions}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Bibliography
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% MEE style
\bibliographystyle{mee}
\bibliography{article}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Appendix section 1}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
